### 2021-03-15 presentations/deer_wolf_CWD.Rmd
Presentation and Cohesivenss fix

---
title: "Can Selective Predation Slow the Spread of CWD?"
subtitle: "Modeling disease, predation, dispersal, and population dynamics in Wisconsin"
author: "Elie Gurarie"
date: "March 30, 2020"
header-includes:	
   - \input{preamble.tex}
fontsize: 11pt
classoption: "aspectratio=43"
output: 
  beamer_presentation:
    theme: "Madrid"
    keep_tex: true
---

```{r setup, include=FALSE}

knitr::knit_hooks$set(mysize = function(before, options, envir) {
  if (before) 
    return(options$size)
})
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, encoding = "utf-8", message = FALSE, warning = FALSE)
```
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
```

# Acknowledgement

\bcol

\col[.3]
\col[.3]
\PlotW{../images/VanDeelen}
\PlotW{../images/FWE}
\col[.3]
\ecol

\bc Prof. Tim Van Deelen \ec

# The setting: **Wisconsin**

\PlotW{../images/Wisconsin_in_United_States.png}

# The setting: **Wisconsin**

  \bc
  \PlotW[.45]{../images/WisconsinVegetation}
  \PlotW[.45]{../images/TensionZone}
  \ec

* **Southwest:** Mainly agricultural / ex-prairie and oak savanna.
* **Northeast:** Conifer-hardwood forest / bogs / also agriculture

# The setting: **Wisconsin**

  \bc  \PlotW[.7]{../images/Treaty1837} \ec
  Wildlife management and fishing rights largely retained by Ojibwe tribes in the North.

# Three characters

\Large

1. White-tailed deer
2. Chronic wasting disease
3. Wolves

# White-tailed deer in Wisconsin

\bcol

\col[.6]

\bbb{\emph{Odoiceulus virginiacus}}

\small
- (over)-abundant: 1.9-2.1 million ind.
- major ecological impacts
- $\approx$ 350,000 hunted annually (and falling)
- \$1.4 billion dollars / year to economy
- a big chunk of which funds research / mitigation of those ecological impacts

\ebb

\col[.4]
\PlotW{../images/deer}
\bc\scriptsize \bl{Todd Hubler - \emph{The Isthmus}}\ec
\ecol

# Chronic Wasting Disease

\bcol

\col[.6]

\small

\bbb{Transmissible spongiform encephalopathy}

- Turns brains into sponges
- Invariably fatal 
- Caused by **prion** 
   - misfolded protein found in nervous system
- Only affects cervids 
   - only TSE in *wildlife* 
- Major focus of concern / research among agencies

\ebb

\col[.4]

\PlotW{../images/misfold}
\PlotW{../images/bse_sponge}

\ecol

# Chronic Wasting Disease

\bcol

\col[.5]

\small

\bbb{Clinical Signs}

"Zombie Disease"

- Emaciation
- Lack of coordination
- Drooping head/ears
- Excessive drooling
- Excessive drinking
- Excessive urination

\ebb

\col[.5]

\PlotW{../images/sickdeer}
\ecol

**Incubation:** (asymptomatic) period lasts on average 18 months

**Transmission:** urine, feces, blood.  Direct contact. Long-term environmental persistence (even uptake by plants).

# Chronic Wasting Disease

\bcol

\col[.4]

\footnotesize

\bbb{Global Expansion}

- 1967 - First detected (mule deer) in research facility Colorado. 
- 1981 - First wild animal (elk) detected in Colorado
- 2002 - Found in wild WTD in Wisconsin
- 2011 - Found in wild WTD in Maryland
- 2017 - Appeared in 3 reindeer in Norway (!) - entire 2000 animal herd summarily executed

\ebb

\col[.6]

\PlotW{../images/CWD_NorthAmerica}

\ecol

# Chronic Wasting Disease: In Wisconsin

Concentrated in southern counties, up to 25\% prevalance.

\bc \PlotW[.6]{../images/CWDprevalence} \ec

***Good data:***  In affected counties, all hunted carcasses need to be tested.  In non-affected counties, a sample of carcasses is tested.

# Wolves in Wisconsin

\small
Extirpated early 1900's.  Re-colonized from Minnesota post-ESA.

\bcol

\col

\PlotW[.9]{../images/WolfMap2019}

\col

```{r WolfNums, fig.width = 4, fig.height = 4}
ww <- read.csv("../data/wolfcounts.csv")
require(ggplot2); require(ggthemes); require(gridExtra)
p1 <- ggplot(ww, aes(Year, Wisconsin)) + geom_point() + ylab("N. wolves in WI")
p2 <- ggplot(ww, aes(Year, Wisconsin)) + geom_point() + geom_smooth() + scale_y_log10()  + ylab("N. wolves in WI")
grid.arrange(p1, p2, ncol = 1)
```

\ecol

Currently, approx. 1000 ind. mainly in North.  Expansion slowing. ***Good Data.***

# The question

\small

Wolves selectively predate on old, young, weak or **infirm(?)** individuals ... though there is no direct evidence w.r.t. CWD (or - actually - other diseases).

Given that CWD is concentrated in the SW - and expanding - and wolves are concentrated in NE - and maybe still expanding? - What happens when they meet?

Specifically, how do wolf presence and selective predation influence:

- **CWD prevalence**
- **CWD spread**
- **Deer abundance**

# Approaches to look at this question

Lots of ***Mathematical Modeling!!***  Mainly, continuous-time, non-spatial SEIR-type ODE's.

\bcol
\col
\PlotW{../images/WildTitle}
\PlotW{../images/WildEquations}

Very influential, but no data (and no spatial structure)

\col
\bc
\PlotW{../images/TannerTitle}
\PlotW[0.7]{../images/TannerEquations}
\ec

Lots of compartments - and some data (but no spatial structure)

\ecol

# Modeling goals

* Capturing dynamics of:
  - **disease**, 
  - **predation**, 
  - **population** 
  - **dispersal**

* Biologically meaningful parameters
  - independently estimated / estimable?

* Provide spatially and temporally explicit predictions

* Balances realism with tractability

# Basic model structure:

## Discrete time / discrete space

- **Annual** - matches data collection and deer biology (birth / seasonal mortality / dispersal?)

- **County-level metapopulation** - matches data reporting and collection

## Two classes: Susceptible and Infected
\small
\begin{align}
S_{i,t+1} &= S_{i,t} - \rd{infected} & + \bl{recruited}& - \rd{died} + \bl{immigrated} - \rd{emigrated} \nonumber\\
I_{i, t+1} &= I_{i,t} + \bl{infected} & &  -\rd{died} + \bl{immigrated} - \rd{emigrated} \nonumber
\end{align}

# Complete model

\bc

\scriptsize
|   |  Susceptible $(S_{i,t+1})$|  Infected $(I_{i,t+1})$|  
|---|:---:|:---:|
| disease          | $- \gamma {S_{i,t} I_{i,t} \over area}$ | $\gamma{ S_{i,t} I_{i,t} \over area}$ |
| predation        | $-\left({S_{i,t} \over S_{i,t}+I_{i,t}}\right) \left({1 \over 1 + \alpha}\right) W_{max}$ |  $-\left({I_{i,t} \over S_{i,t}+I_{i,t}}\right) \left({\alpha \over 1 + \alpha}\right) W_{max}$      |
| other mortality  | $- \mu_s S_{i,t}$ |  $- \mu_I I_{i,t}$ |
| recruitment      |  $\rho S_{i,t}(1 - S_{i,t}/K_i)$  |  |
| immigration      | $\sum_{j} M_{S,ij}$  | $\sum_{j} M_{I,ij}$ |
| emigration       | $-\sum_{j} E_{s,ji}$ | $-\sum_{j} E_{i,ji}$ |

\PlotW[.7]{../images/gif/crudity_1}

\ec

# Recruitment

##
Approximate number of new (non-infected) individuals entering population annually:
\Large
$$\rho \, S_{i,t}(1 - S_{i,t}/K_i)$$
\normalsize

Assumptions:

1. Logistic, density dependent growth with carrying capacity: $K$
2. Intrinsic adult recruitment rate, $\rho$, uniform across range

# Mortality

## Non-Wolf:

Annual mortality rate: \rd{$\mu_S$} and \rd{$\mu_I$ }:

- Equal (!?)
- Constant across range (?))

# Mortality

## Wolf:

Assume: **Wolves gonna kill what wolves gonna kill** - (almost entirely deer).

- \small e.g. 10 wolves, each kills 20 deer / year = 200 deer/year/county killed

\normalsize Apportioning of S and I killed is proportional to the *availability* ($I \over S+I$) and the *preference* ($\alpha \over 1+\alpha$).

- \small $P(I \, killed) \propto P(I) \times P(killed \,|\, I) = \left({I \over S + I}\right)  \left(\alpha  \over 1 + \alpha \right) N_w K_{rate}$
- \small  $P(S \, killed) \propto P(S) \times P(killed \,|\, S) = \left({S \over S + I}\right)  \left(1  \over 1 + \alpha \right) N_w K_{rate}$

where \gr{$\alpha$} is the odds of selecting infected over non-infected deer, i.e. the \gr{Predation Selection Coefficient} (presumably $>$ 1). 
\nd

__This is the key, unobservable parameter of interest!__

# Disease dynamics

## Number of infected ...
(by county by year)
\Large
$$- \gamma {S \times I \over area}$$
\large

- $\gamma$ is infection rate per encounter  
- Total number of infections is higher if areas are smaller (densities are higher). 
- Lots of mechanism and structure swept under rug here!

# Dispersal

**Necessary for disease to spread! (or not spread)**

## Immigration
$$M_i  = {\kappa_m} \sum_{j} S_j \exp {\left(- {d_{ij} \over \lambda} \right)^\beta}$$

## Emigration
$$E_j  = {\kappa_e} S_i \sum_{j} \exp {\left(- {d_{ij} \over \lambda} \right)^\beta}$$

\footnotesize
- **Exponential** ($\beta = 1$) or **Gaussian** ($\beta = 2$) dispersal kernel with mean $\lambda$
- $\kappa$ - normalization, so that probability of emigrating from site $j$ sums to 1
- Assumes homogenous dispersal in all directions, dependent only on distance. 
- *Estimable (and refinable) from movement data!?*

# Data

## **Deer Abundance**
Wisconsin DNR winter population survey: [https://dnr.wi.gov/topic/hunt/maps.html](https://dnr.wi.gov/topic/hunt/maps.html)
\bc
\PlotW[.45]{../images/overwinter.pdf}
\PlotW[.45]{../images/winterpoppertotal.pdf}
\ec
Fall population estimates - total harvest, by county.

\nd Working assumption: Carrying Capacity $K_i = 2 N_i$.

# Data

## **CWD prevalence**
Wisconsin DNR CWD monitoring efforts (by county) 
\nd

\bc
\PlotW[.7]{../images/CWDprevalence}
\ec

\small https://dnr.wi.gov/wmcwd/Summary/YearCounty/2019

# Data

## **Wolves**

Latest estimate from DNR: 950 ind.

\bc \PlotH[.5]{../images/WolfMap2019.png} \,\,\, \PlotH[.5]{../images/WolfMapElie.png} \ec

County data not readily available ... so I allocated 1000 wolves across the counties north of this line.

# Model

## Shiny App.

A brand new skill!
Facilitates exploring parameters and visualizing results.

\bc
(enjoy demo)
\ec

# A Result:  Selective Predation Decreases CWD Prevalence!
\bc
\PlotW[.9]{../images/SimResults1.png}
\ec

\small
In ALL parameterizations, wolves depress CWD.  Note - dispersal scale (10 and 80 km) AND shape both important.

\footnotesize `rho = 0.5, gamma = .02, mu_S = 0.06, mu_I = 0.06, W_max = 60, lambda = 10 or 80`

# Next steps

\footnotesize

## Model structure
* Add **Male / Female** sex classes!
* Separate **Infected / Asymptomatic** from **Infected / Symptomatic**
* Assess assumptions:  Density dependence? | Disease transmission? 
\pause

## Data
- Obtain better **wolf distribution** and **predation** data
- Use **Harvest** for mortality!
- Use **GPS data** for dispersal portion
- Fit to historical data!?
  - Infer $\gamma$ by matching to observed CWD spread? 
\pause

## Larger strategy
- Thoroughly analyze / explore parameter space
- Find PhD student to do the work!?
- Get funding!

#

\bc
\Huge
Thanks!
\ec