### 2021-03-25 presentations/memorymigration_v2.Rmd
Presentation more or less good to go

---
title: "Modeling tactical and memory-driven migration in dynamic environments"
author: "Sriya Potluri and Elie Gurarie"
date: "3/24/2021"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE)
```

# Premise
<div class='left' style='float:left;width:48%'>
To optimally forage in a dynamic environment, you need both **tactical** responses (adapting to local changes in resource gradients), and **strategic** responses (using memory to guide long-term decisions).

</div>
<div class='right' style='float:right;width:48%'>
**Tactics**

![](images/tacticsim.png){width=100%}

**Memory**

![](images/memorysim.png){width=100%}
</div>

# Model: Diffusion advection

$${\frac{\partial u(x,t)}{\partial t}} = -\varepsilon {\frac{\partial^2 u(x,t)}{\partial x^2}} + \alpha \frac{\partial}{\partial x}\left(u \frac{\partial h(x,t)}{\partial x}\right) + \sum_{i = 0}^n \beta_i \frac{\partial u(x, t - i\tau)}{\partial x}$$

- $\epsilon$: diffusion, exploration / search
- $\alpha$: taxis, resource following
- $\beta_0$: cohesion, tendency to stick to group
- $\beta_1$: memory, tendency to do what was done previous year

Builds on series of projects exploring "cognitive" variations chemotaxis (e.g. Fagan et al. 2017, 2019).

# Seasonal resource

$$h(x,t, \theta) = \chi B(x/\chi, a(t, \theta), b(t, \theta))$$

<div class='left' style='float:left;width:48%'>
1. The total amount constant.
2. Between seasons, uniformly distributed. 
3. Resource concentrates symmetrically in time and space

Importantly: we explore a range of **spatial concentrations** and **temporal durations**.

</div>

<div class='right' style='float:right;width:48%'>
![](images/exampleResourcePlots-1.png){width=100%}
</div>

---

<div class='left' style='float:left;width:48%'>
## Running the model

- R package `deSolve` and `ReacTran` - fluid transport pde solvers.
- Functions all in `memorymigration` GitHub project. 
- SESYNC cluster 
- server-based Rstudio
</div>

<div class='right' style='float:right;width:48%'>
## Assessing the model

Four metrics:

1. Social cohesiveness
2. Foraging Efficiency
3. Migratoriness
4. Time to equilibrium

</div>

# Metrics: Cohesiveness

<div class='left' style='float:left;width:50%'>

$$SC = 1-\frac{\sqrt{12} \, }{X_{max}}\overline{\sigma}$$

$$\overline{\sigma} = {1\over \tau} \int \sigma_u(x,t) \,dt$$
where $$\sigma_u^2(t) = \int_D \left(x - \mu_u(t)\right)^2 \, u(x,t)\, dx$$

and $\mu_u$ is the mean of $u$ at time $t$.

Definition: The mean standard deviation of the population across the equilibrium year

</div>

<div class='right' style='float:right;width:50%'>

```{r SC_example, fig.width = 6}
require(memorymigration)
data(resources)
par(mfrow = c(2,1), mar = c(3,3,3,1), mgp = c(2,.25,0), las = 1, tck = 0.01)

image(1:100,1:100,resource_R2, main = "SC = 0.36")
contour(1:100,1:100,resource_R2, add = TRUE)
#computeCohesiveness(resource_R2, world)

image(1:100,1:100,resource_R3, main = "SC = 0.11")
contour(1:100,1:100,resource_R3, add = TRUE)

#computeCohesiveness(resource_R3, world)

```

</div>

# Social Cohesiveness: Problem

For skewed of bimodal population distribution, standard deviation are a poor measure ...
$\epsilon = 8; \alpha = 200; \beta_0 = 500; \beta_1 = 0$

```{r badSCexample, cache = TRUE, fig.width = 10, eval = FALSE}
load("images/Sim1.rda")
png("images/BadSC.png", width = 2400, height = 1200, res = 300)
par(mfrow = c(1,2), bty  = "l", mgp = c(1.5,.5,0), tck = 0.01)
require(fields)
computeCohesiveness(sim[[length(sim)]], world)
image.plot(sim[[length(sim)]], main = "looks cohesive, but SC = 0.01")
snapshot <- sim[[length(sim)]][75,]
plot(snapshot, type = "l", main = "because bimodal (t = 75).", ylab = "")
dev.off()
```

![](images/BadSC.png){width=100%}

**Not sure how to deal with this ...**

# Metrics: Foraging Efficiency

$$FE = {1\over \tau} \int_{0}^\tau \int_{0}^{X_{max}} \sqrt{u(x,t) \, h(x,t)} \, dx\,dt$$

**Bhattacharyya Coefficient** quantifies similarity between two distributions.

```{r, eval = FALSE}
load("images/Sim1.rda")
png("images/FEexample.png", width = 2400, height = 1200, res = 300)
par(mfrow = c(1,2), bty  = "l", mgp = c(1.5,.5,0), tck = 0.01)
require(fields)
image.plot(1:100, 1:100, resource_R2, main = "Resource", xlab = "time", ylab = "space")
image.plot(1:100, 1:100, sim[[length(sim)]], main = "Population: FE = 0.82", xlab = "time")
dev.off()

computeEfficiency(resource_R2, sim[[length(sim)]], world)
```

<center>
$\epsilon = 8; \alpha = 200; \beta_0 = 500; \beta_1 = 0$

![](images/FEexample.png){width=70%}
</center>

# Metrics: Foraging Efficiency - Issues

Seems to always be very high!

Also - assumes that there is some sort of space inhibition ... the actually highest efficiency is when animals pile in at indefinitely high density at the location of highest resource availability.

<center>

![](images/FEexample.png){width=70%}

</center>

# Metrics: Migratoriness

$$MI = 1-min(O(t_1, t_2))$$
$$O(t_1, t_2) = \int_D \sqrt{u(x, t_1) \, u(x, t_2)} \, dx $$

Definition: MINIMUM OVERLAP of population at all possible pairs of times within year.

```{r, eval = FALSE}
load("images/Sim1.rda")
png("images/MIexample.png", width = 2400, height = 1200, res = 300)
par(mfrow = c(1,2), bty  = "l", mgp = c(1.5,.5,0), tck = 0.01)
require(fields)
image.plot(1:100, 1:100, resource_R2, main = "Resource", xlab = "time", ylab = "space")
image.plot(1:100, 1:100, sim[[length(sim)]], main = "Population: MI = 0.07", xlab = "time")
dev.off()

computeEfficiency(resource_R2, sim[[length(sim)]], world)
```

<center>

$\epsilon = 8; \alpha = 200; \beta_0 = 500; \beta_1 = 0$

![](images/MIexample.png){width=70%}

</center>

# Questions that this model can answer:

<div class='left' style='float:left;width:60%'>

Q1. How does **memory** and **sociality** influence **migratoriness** and **foraging efficiency** under different **resource dynamics**?

Q2. How long does it take for migration to emerge from a non-migratory initial condition?

Q3. As conditions CHANGE, how resilient is the population at various Diffusion, Advection, Cohesion, Memory values?  
- spatial drift
- temporal drift
- increased variability

</div>

<div class='right' style='float:right;width:40%'>

![](images/brainstorm.png){width=100%}

</div>

# Runs

Lots of them! (8100)

## Model parameters

- $\alpha$ - tactic response: 0, 100, 1000
- $\epsilon$ - diffusion: 10, 30, 100
- $\beta_0$ - sociality: 0, 100, 200, 300, 400, 500
- $\beta_1$ - memory: 0, 100, 200, 300, 400, 500

## Resource pulse parameters

- Duration (2 x sd): 6,12,18,24,30
- Spatial extent (2 x sd): 6,12,18,24,30

# Basic Metrics

<div class='left' style='float:left;width:60%'>
actic response**
2. High overall mobility (high tactic + high diffusion) =  high foraging efficiency
3. Slow diffusion leads to longer time to attain equilibrium

We mainly focus on **low diffusion** and **high tactic**, where most of the (migratory) "action" is.

</div>

<div class='right' style='float:right;width:40%'>
![](images/MetricComparisons.png){width=100%}
</div>

# Example simulations

![](images/resource.png){width=20%}![](images/sim0.png){width=70%}

# Example: Memory 0 / Sociality 500

![](images/resource.png){width=20%}![](images/sim2.png){width=70%}

# Example: Memory 500 / Sociality 0

![](images/resource.png){width=20%}![](images/sim1.png){width=70%}

# Example: Memory 200 / Sociality 200

![](images/resource.png){width=20%}![](images/sim3.png){width=70%}

# Example: Memory 500 / Sociality 200 (sometimes things fall apart)

![](images/resource.png){width=20%}![](images/sim4.png){width=70%}

# Some results

<div class='left' style='float:left;width:50%'>

1. Migratoriness results from **concentrated** and **long** resource pulses. 
2. Both migratoriness and sociality **increase migratoriness**. 
3. Foraging efficiency is highest for **diluted** and **short** resource pulses (resource is distributed mainly uniformly).
4. In a "stationary" (equilibrium) scenario, memory and sociality don't contribute much to foraging efficiency (**caveated!**).

</div>

<div class='right' style='float:right;width:50%'>

![](images/ComparingResourceDynamics.png){width=100%}

</div>

---

## Main take-away

1. This is a convenient, flexible and fast tool for exploring a wide range of memory / tactic / sociality dynamics

## Going forward

"Stress" the system to assess plasticity of a population:

- Make the pulses more isolated (weaken usefulness of tactical strategy).
- Drift the **location** and **timing** of pulse
- Make the pulse more **patchy** and **stochastic**