### 2021-04-07 code/IslandResourcesExploration.Rmd

---
title: "Island Resources Exploration"
author: "E. Gurarie, S. Potluri, B. Fagan, others"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
theme: cerulean
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE)
```

```{r, message=FALSE, warning=FALSE}
require("memorymigration")
```

## Gaussian Initial Conditions; Resource t3 x3

```{r, message=FALSE, warning=FALSE, eval=FALSE}

  data(world)
  data(resources_island)
  world_gaussian$resource <- resources_island[["R_t3_x3"]]
  image(world_gaussian$resource)

  sim0 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 0, beta1 = 0), verbose = TRUE)
  sim1 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 0, beta1 = 500), verbose = TRUE)
  sim2 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 500, beta1 = 0), verbose = TRUE)
  sim3 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 250, beta1 = 250), verbose = TRUE)
  sim4 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 500, beta1 = 200), verbose = TRUE)

  save(sim0, sim1, sim2, sim3, sim4, file = "results/islandresources/Gaussianalpha1000epsilon10_R_t3_x3.rda")
} else load("results/islandresources/Gaussianalpha1000epsilon10_R_t3_x3.rda")

png("presentations/images/gaussiantactict3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim0)) image(1:100, 1:100, sim0[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only tactic response", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/gaussianmemoryt3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim1)) image(1:100, 1:100, sim1[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only memory", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/gaussiansocialityt3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim2)) image(1:100, 1:100, sim2[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only sociality", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/gaussianmemorysociality1t3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim3)) image(1:100, 1:100, sim2[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("With memory and sociality beta1=beta0=250", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/gaussianmemorysociality2t3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(2,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:14) 
  image(1:100, 1:100, sim4[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("With memory and sociality beta1=200, beta0=500", outer = TRUE, cex.main = 2)
dev.off()

```

![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussiantactict3x3.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussianmemoryt3x3.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussiansocialityt3x3.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussianmemorysociality1t3x3.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussianmemorysociality2t3x3.png)

## Gaussian Initial Conditions; Resource t15 x15

```{r, message=FALSE, warning=FALSE, eval=FALSE}

  data(world)
  data(resources_island)
  world_gaussian$resource <- resources_island[["R_t15_x15"]]
  image(world_gaussian$resource)

  sim0 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 0, beta1 = 0), verbose = TRUE)
  sim1 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 0, beta1 = 500), verbose = TRUE)
  sim2 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 500, beta1 = 0), verbose = TRUE)
  sim3 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 250, beta1 = 250), verbose = TRUE)
  sim4 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 500, beta1 = 200), verbose = TRUE)

  save(sim0, sim1, sim2, sim3, sim4, file = "results/islandresources/Gaussianalpha1000epsilon10_R_t15_x15.rda")
} else load("results/islandresources/Gaussianalpha1000epsilon10_R_t15_x15.rda")

png("presentations/images/gaussiantactict15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim0)) image(1:100, 1:100, sim0[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only tactic response", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/gaussianmemoryt15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim1)) image(1:100, 1:100, sim1[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only memory", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/gaussiansocialityt15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim2)) image(1:100, 1:100, sim2[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only sociality", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/gaussianmemorysociality1t15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim3)) image(1:100, 1:100, sim3[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("With memory and sociality beta1=beta0=250", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/gaussianmemorysociality2t15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(4,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim4)) image(1:100, 1:100, sim4[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("With memory and sociality beta1=200, beta0=500", outer = TRUE, cex.main = 2)
dev.off()
```

![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussiantactict15x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussianmemoryt15x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussiansocialityt15x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussianmemorysociality1t15x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/gaussianmemorysociality2t15x15.png)

## Sinusoidal Initial Conditions; Resource t3 x3

```{r, message=FALSE, warning=FALSE, eval=FALSE}

  data(world)
  data(resources_island)
  world_sinusoidal$resource <- resources_island[["R_t3_x3"]]
  image(world_sinusoidal$resource)

  sim0 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 0, beta1 = 0), verbose = TRUE)
  sim1 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 0, beta1 = 500), verbose = TRUE)
  sim2 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 500, beta1 = 0), verbose = TRUE)
  sim3 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 250, beta1 = 250), verbose = TRUE)
  sim4 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 500, beta1 = 200), verbose = TRUE)

  save(sim0, sim1, sim2, sim3, sim4, file = "results/islandresources/Sinusoidalalpha1000epsilon10_R_t3_x3.rda")
} else load("results/islandresources/Sinusoidalalpha1000epsilon10_R_t3_x3.rda")

png("presentations/images/sinusoidaltactict3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim0)) image(1:100, 1:100, sim0[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only tactic response", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/sinusoidalmemoryt3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim1)) image(1:100, 1:100, sim1[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only memory", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/sinusoidalsocialityt3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim2)) image(1:100, 1:100, sim2[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only sociality", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/sinusoidalmemorysociality1t3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim3)) image(1:100, 1:100, sim3[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("With memory and sociality beta1=beta0=250", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/sinusoidalmemorysociality2t3x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(2,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim4)) image(1:100, 1:100, sim4[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("With memory and sociality beta1=200, beta0=500", outer = TRUE, cex.main = 2)
dev.off()

```
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidaltactict3x3.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidalmemoryt3x3.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidalsocialityt3x3.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidalmemorysociality1t3x3.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidalmemorysociality2t3x3.png)

## Sinusoidal Initial Conditions; Resource t15 x15

```{r, message=FALSE, warning=FALSE, eval=FALSE}

  data(world)
  data(resources_island)
  world_sinusoidal$resource <- resources_island[["R_t15_x15"]]
  image(world_sinusoidal$resource)

  sim0 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 0, beta1 = 0), verbose = TRUE)
  sim1 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 0, beta1 = 500), verbose = TRUE)
  sim2 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 500, beta1 = 0), verbose = TRUE)
  sim3 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 250, beta1 = 250), verbose = TRUE)
  sim4 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 500, beta1 = 200), verbose = TRUE)

  save(sim0, sim1, sim2, sim3, sim4, file = "results/islandresources/Sinusoidalalpha1000epsilon10_R_t15_x15.rda")
} else load("results/islandresources/Sinusoidalalpha1000epsilon10_R_t15_x15.rda")

png("presentations/images/sinusoidaltactict15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim0)) image(1:100, 1:100, sim0[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only tactic response", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/sinusoidalmemoryt15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim1)) image(1:100, 1:100, sim1[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only memory", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/sinusoidalsocialityt15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,4), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim2)) image(1:100, 1:100, sim2[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("Only sociality", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/sinusoidalmemorysociality1t15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(1,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim3)) image(1:100, 1:100, sim3[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("With memory and sociality beta1=beta0=250", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/sinusoidalmemorysociality2t15x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(4,6), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:length(sim4)) image(1:100, 1:100, sim4[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("With memory and sociality beta1=200, beta0=500", outer = TRUE, cex.main = 2)
dev.off()
```
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidaltactict15x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidalmemoryt15x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidalsocialityt15x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidalmemorysociality1t15x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/sinusoidalmemorysociality2t15x15.png)

## 60 year runs Gaussian. 
```{r, message=FALSE, warning=FALSE, eval=FALSE}
  data(world)
  data(resources_island)

  world_gaussian$resource <- resources_island[["R_t12_x12"]]
  sim0 <- runManyYears(world_gaussian, parameters = c(epsilon = 5, alpha = 1000, beta0 = 100, beta1 = 100), verbose = TRUE)

  world_gaussian$resource <- resources_island[["R_t3_x15"]]
  sim1 <- runManyYears(world_gaussian, parameters = c(epsilon = 10, alpha = 1000, beta0 = 300, beta1 = 300), verbose = TRUE)

  world_gaussian$resource <- resources_island[["R_t15_x3"]]
  sim2 <- runManyYears(world_gaussian, parameters = c(epsilon = 5, alpha = 2000, beta0 = 300, beta1 = 200), verbose = TRUE)

  world_gaussian$resource <- resources_island[["R_t6_x3"]]
  sim3 <- runManyYears(world_gaussian, parameters = c(epsilon = 5, alpha = 500, beta0 = 200, beta1 = 200), verbose = TRUE)

  save(sim0, sim1, sim2, sim3, file = "results/islandresources/60yearruns_gaussianisland.rda")
} else load("results/islandresources/60yearruns_gaussianisland.rda")

png("presentations/images/60yearrungaussiant12x12.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(3,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:15) 
  image(1:100, 1:100, sim0[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("T12X12, alpha=1000, epsilon=10, beta1=beta0=100", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/60yearrungaussiant3x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(3,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:15) 
  image(1:100, 1:100, sim1[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("T3X15, alpha=1000, epsilon=10, beta1=beta0=300", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/60yearrungaussiant15x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(3,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:15) 
  image(1:100, 1:100, sim2[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("T15X3, ealpha=2000, epsilon=10, beta0=300, beta1=200", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/60yearrungaussiant6x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(3,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:15) 
  image(1:100, 1:100, sim3[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("T6X3, epsilon=500, alpha=5, beta1=beta0=200", outer = TRUE, cex.main = 2)
dev.off()

```
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/60yearrungaussiant12x12.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/60yearrungaussiant3x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/60yearrungaussiant15x3.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/60yearrungaussiant6x3.png)

## 60 year runs Sinusoidal. No matter the initial condition, all of the 60 years runs happen with the relatively the same parameter combinations. There are less 60 year runs with sinusoidal intial condition compared to gaussian initial condition. 
```{r, message=FALSE, warning=FALSE, eval=FALSE}

  data(world)
  data(resources_island)

  world_sinusoidal$resource <- resources_island[["R_t12_x12"]]
  sim0 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 5, alpha = 1000, beta0 = 100, beta1 = 100), verbose = TRUE)

  world_sinusoidal$resource <- resources_island[["R_t3_x15"]]
  sim1 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 10, alpha = 1000, beta0 = 300, beta1 = 300), verbose = TRUE)

  world_sinusoidal$resource <- resources_island[["R_t15_x3"]]
  sim2 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 5, alpha = 2000, beta0 = 300, beta1 = 200), verbose = TRUE)

  world_sinusoidal$resource <- resources_island[["R_t6_x3"]]
  sim3 <- runManyYears(world_sinusoidal, parameters = c(epsilon = 5, alpha = 500, beta0 = 200, beta1 = 200), verbose = TRUE)

  save(sim0, sim1, sim2, sim3, file = "results/islandresources/60yearruns_sinusoidalisland.rda")
} else load("results/islandresources/60yearruns_sinusoidalisland.rda")

png("presentations/images/60yearrunsinusoidalt12x12.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(3,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:15) 
  image(1:100, 1:100, sim0[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("T12X12, alpha=1000, epsilon=10, beta1=beta0=100", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/60yearrunsinusoidalt3x15.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(3,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:15) 
  image(1:100, 1:100, sim1[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("T3X15, alpha=1000, epsilon=10, beta1=beta0=300", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/60yearruntsinusoidal15x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(3,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:15) 
  image(1:100, 1:100, sim2[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("T15X3, alpha=2000, epsilon=10, beta1=200, beta0=300", outer = TRUE, cex.main = 2)
dev.off()

png("presentations/images/60yearrunsinusoidalt6x3.png", width = 3000, height = 1500, res = 300)
par(mfrow = c(3,5), mar = c(0,0,3,0), oma = c(2,2,4,2), tck = 0.01)
for(i in 1:15) 
  image(1:100, 1:100, sim3[[i]], main = paste("year", i-1), yaxt = "n", xaxt = "n")
title("T6X3, alpha=500, epsilon=5, beta1=beta0=200", outer = TRUE, cex.main = 2)
dev.off()

```
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/60yearrunsinusoidalt12x12.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/60yearrunsinusoidalt3x15.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/60yearrunsinusoidalt15x6.png)
![](/Users/sriyapotluri/Documents/Research/memorymigration/presentations/images/60yearrunsinusoidalt6x3.png)

## NA Runs There were two runs that had missing values of foraging efficiency and the final similatiry in the sinusoidal intitial condition. 
```{r, message=FALSE, warning=FALSE, results='asis'}

load("/Users/sriyapotluri/Documents/Research/memorymigration/results/Results_Island_Gaussian.rda")
r.gaussian <- results
r.gaussian[!complete.cases(r.gaussian),]

load("/Users/sriyapotluri/Documents/Research/memorymigration/results/Results_Island_Sinusoidal.rda")
r.sinusoidal <- results
naruns <- r.sinusoidal[!complete.cases(r.sinusoidal),]

library(knitr)
kable(naruns, caption="NaN runs")
```

```