### 2021-06-14 text/ms.Rmd
...

---
title: "Modeling memory and taxis to analyze migrations"
author: "E. Gurarie, S. Potluri, B. Fagan, others"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
theme: cerulean
editor_options:
  chunk_output_type: inline
bibliography: ../library/memorymigration.bib
csl: ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, eval = FALSE)
```

# Introduction

Seasonal migrations are extremely widespread among terrestrial, aquatic, avian and invertebrate species [@Dingle2014].  For many species, migration is an extremely successful strategy, allowing a far greater number of individuals to inhabit landscape which might not otherwise be able to support large numbers year round [@Fryxell1988]. The ultimate evolutionary cause of seasonal migration, essentially, is that the energetic and survival related costs of migration are outweighed by the fitness benefits of accessing suitable seasonal resources, whether those are for energetic gain, predator avoidance, a suitable physical, biotic or social environment for reproduction [@Avgar2014].

Much as the ultimate causes of migration vary across species, the proximate causes, drivers and mechanism can vary considerably across and even within species [@Shaw2016]. Some migrants follow a "green wave" of spring vegetation as it flowers across altitudinal or latitudinal gradients [@Bischof2012;@Kolzsch2015]. These migrations can be considered "tactical", as they can occur - as an extreme simplification - purely as  response to local conditions.   Other migrants perform long-distance migrations in anticipation that critical resources will be available at the time of arrival at the end point of migration (**refs**).  This second behavior involves the greatest trade-off between the costs of migration against the benefits of accessing resources, whether food, suitable habitat for breeding, or predator refuge, that are highly seasonal and localized (**refs**).  This approach can be considered "strategic" in the sense that it is driven not by immediate cues but by a remembered anticipation [@Bracis2017].

Migratory species are often considered to be more vulnerable to environmental change, as a disruption in either of the seasonal ranges or along a migratory corridor can have significant negative impacts [@Wilcove2008;@Kauffman2021].  On the other hand, it has been argued that migratory species might be more resilient to disruptions due to their wide-ranging mobility [@Robinson2009]. Clearly, the ability to perform a migration without local cues in only possible if the behavior is hard-cued or remembered.  On the other hand, a strictly programmed behavior can be maladaptive if conditions change.

Diffusion-advection models have a long pedigree in animal movement modeling [@Skellam1951;@Turchin1998;@Okubo2001]. These models are grounded in the general idea that animal movements - somewhat like movements of physical particles - combine a random (diffusive) component with a directed (advective) component. While direct relationships between diffusion models and movement data are somewhat tenuous [@Gurarie2011a], as a theoretical tool for exploring processes they are invaluable for their versatility and the relative ease of numeric computation of the partial differential equations (PDEs) that are used to describe diffusion-advection models mathematically. Thus, much theoretical and some applied work has been done on refining the basic assumptions of diffusion models, e.g. by including heterogeneity in populations [@Skalski2003;@Gurarie2009], fat-tailed dispersal kernels [@Kot1996], non-linear or otherwise complex responses to resources and consepecifics (**refs**).

Perhaps the greatest difference between animals and randomly moving passive particles described by diffusion models is cognition, including social behavior and memory. Refinements to diffusion-advection equations have revealed conditions under which non-local information gathering [@Fagan2017] and behavioral switching may confer foraging advantages [@Fagan2019], in particular when resources are dynamic and patchy. The interacting role of memory and sociality, in contrast, have been comparatively little studied.

Here, we develop a diffusion-advection model with memory to explore the resilience of a migratory population under various dynamic, seasonal resource distributions.  In formulating the model, our goal is to identify the minimum set of movement and memory parameters required to generate an adaptive, migratory behavior. This includes the ability to learn to migrate from a non-migratory initial conditions, simulating the release of naive animals in a seasonal environment [@Jesmer2018], or to lose the propensity to migrate if the resource distribution does not require it, also a commonly observed phenomenon (**refs**).  Our ultimate goal is to assess the role of long and short-term memory in the resilience (or fragility) of a migratory population against changing patterns in resource distribution that are the consequences of rapid climate change.

We anticipate that under many conditions a blending of *tactical* (i.e. direct response to resource availability or perception) and *strategic* (i.e. memory-driven and forward-thinking) behavior help foragers navigate dynamic, seasonal environments.  Over-reliance on either strategy should be maladaptive.  We further anticipate that a a shorter-term memory updating is needed to navigate trends in resource spatial distribution and temporal distribution (phenology), but that a longer-term reference memory is needed to navigate resource distributions that are increasingly stochastic.

# Methods

## Memory movement model

In designing our study, our goal was to develop a minimal heuristic in which the following processes were explicitly modeled: (1) Random or exploratory movement, (2) attraction to resources, (3) a long-term (or *reference*) memory of large-scale movement behavior, (4) a short-term (or *working*) memory that updates movement behavior based on recent experience, (5) some social aspect to the learned behavior.  
A diffusion-advection equation provided a computationally efficient and versatile framework for examining just such a system.  We consider a population moving in one dimension in a constrained domain ${\cal D}$ and distributing itself according to the following equation:

$${\frac{\partial u}{\partial t}} = -\varepsilon {\frac{\partial^2 u}{\partial x^2}} + 
\alpha \frac{\partial}{\partial x}\left(u \frac{\partial h}{\partial x}\right) + 
\beta \frac{\partial}{\partial x}\left(v_s(u)\right) + 
v_m(t)
$$

where  $u$ represents the population distributed in time and space.  The first term is the diffusion term, capturing the fast time-scale exploration and "random" movements of individuals, with $\varepsilon$ is the diffusion rate.

The second term represents the attraction to a dynamic resource $h$, with the proportionality of the advection to the gradient of the resource given by the parameter $\alpha$ (note, the population and resource distributions are functions of both space and time $u(x,t)$ and $h(x,t)$ - we omit the dependent variables in the notation for brevity).  This is the well-studied standard chemotaxic resource-following behavior. We borrow the general notation from our earlier related work [@Fagan2017;@Fagan2019].

The third term captures the collective or social advection term of the population via a non-local, density dependent function $v_s(u,x)$.  If this function takes the form of a convolution around a non-local kernel $k$, i.e.~
$$v_s(u) = k(x) * u(x)$$
and that kernel is odd, an attractive or "swarming" behavior can be generated [@Mogilner1999]. We use the kernel analyzed by @Mogilner1999: 
$$k(x) = \frac{x}{2\lambda^2} \exp(-x^2/2\lambda^2).$$ 
The convolution of $u$ with this kernel has the property of pushing the population in a positive direction when $x < \widehat{u}$, and in a negative direction when $x > \widehat{u}$.  The parameter $\lambda$ is a length scale of sociality - roughly one-half the size of the "swarm", which $\beta$ is a parameter which quantifies the overall strength of sociality.

Finally, the last term captures the direct advection that emerges from a memory-driven migratory behavior. This term evolves with a set of parameters $\theta_y$ that slowly change each year $y \in \{0,1,2,...\}$, i.e. the count of periods $\tau$: $y = \lfloor t/\tau \rfloor$.

The migration speed is specified by six parameters $\theta$: the timing of the start and duration of two anticipated seasons (nominally, winter and summer) $t_1$, $dt_1$, $t_2$, $dt_2$, and the spatial coordinates of the population centroid for each season $x_1$ and $x_2$.  Thus the remembered migratory speed term is a simple step function given by:

$$
v_m(t, \theta_y) = \begin{cases}
0 &;  & t > t_1 &\text{and} & t \leq t_1 + dt_1 \\ 
s_{12} &;  &  t > t_1 + dt_1 &\text{and} & t \leq t_2 \\ 
0 &; & t> t_2 &\text{and} &t \leq t_2 + dt_2 \\ 
s_{21} &; &  t > t_2+dt_2 & \text{or} & t \leq t_1\\ 
\end{cases}
$$

  where the migration speeds $s_{12}$ and $s_{21}$ from the respective ranges are set such that they arrive at $x_1$ at $t_1$, leave at $t = t_1 + dt_1$, arrive at $x_2$ at $t = t_2$, i.e. $s_{12} = \frac{x_2-x_1}{t_2 - (t_1 + dt_1)}$ and $s_{21} =  \frac{x_1-x_2}{t_1 - (t_2 - \tau + dt_2)}$.  This  step-like migration function is a one-dimensional version of the migration parameters estimated individually [@Gurarie2017] and for populations [@Gurarie2019] in empirical studies.

We consider these six parameters to be the known or remembered determinants of the migratory behavior, with an initial set $\theta_0$ determining the reference migration behavior. This reference migration is updated each year by the experience of the population. To perform this updating, we estimate a new set of parameters $\widehat{\theta_y}$ after each year, and combine these new parameters with the reference parameters according to the following weighted mean:

$$\theta_{y+1} = \kappa^y \, \theta_o + \left(1-\kappa^y\right)\,\widehat{\theta_y}$$
where each of the six parameters is updated identically. The estimates $\widehat{\theta_y}$ are obtained via a least-squares minimization of the migration track ($m(t,\theta) = \int_0^t v_m(t',\theta_y) \, dt'$) against the spatial mean of the population process in year $y$ (i.e. $\widehat{u}(t) = \int_X u_y(t, x) dx$).  The parameter $\kappa \in (0,1)$ captures the reliance on that long-term memory. When $\kappa = 0$, all of the actionable memory is from the preceding year.  When $\kappa = 1$, the actionable memory is entirely the reference memory.

The model is confined to a one-dimensional bounded domain $[-\chi,\chi]$, with no flux outside of the boundaries i.e. ${\frac{\partial u(-\chi,t)}{\partial t}} = {\frac{\partial u(\chi,t)}{\partial t}} =0$.  As there are no birth or death processes, the total population remains fixed and constant, for convenience integrating to 1.  Furthermore, the parameters remain constant throughout time, with no adaptation or mutation-selection process.  Our interest is entirely in the ability of a fixed set of memory and movement parameters to ``navigate'' an intra- and interannually dynamic, seasonal environment.

<!-- A few special cases of this model have been studied in the past. @Shigesada1979 (section 3) thoroughly analyzed the resource following diffusion model with no memory and a static resource (i.e.~$\beta = 0$ and $h(x,t) = h(x)$), showing that at equilibrium, the log of the population density is proportional to $h(x)/\alpha$.  @Shi2019 consider a single period memory model ($\kappa = 0$) in detail, with the addition of an independent birth-death process. Importantly, the authors prove the well-posedness (existence, uniqueness, and positivity) of this model, with arguments that are equally valid for our model. @Fagan2017 numerically explored an extension of this model with no memory but a dynamic resource and the addition of non-local information, i.e.~a model where the resource $h(x,t)$ is "perceived" via a convolution with a detection function. @Fagan2019 added a behavioral switching behavior to the non-local information model.  The essential innovation of our study is the inclusion of an explicit, slowly-varying migratory advection term.

**Figure 1.** *Illustration of memory-based velocity term as a logistic function of space (x-axis), mean memory ($\widehat{u_m}(t)$ - vertical dashed line) and mean movement rate ($d\widehat{u_m}/dt$ - horizontal dashed line). Various values of the social spatial scale parameter $\beta$ reflects the maximum slope of the logistic curve such that at small values of the spatial scale, there is a steep difference in the speed around the mean location (blue curve), and at large values of the spatial scale, there as a weak response of the speed to the mean location.  In all cases, at locations beyond the mean, the speed is lower than at locations lagging behind the mean.*
-->

## Seasonal resource

We ran our numerical experiments on a spatial domain $x \in [-100,100]$, and a periodicity $\tau = 100$ (i.e.~100 day years).  We were interested in an approximately periodic resource dynamic, i.e. one in which $h(x,t \approx h(x, t-\tau))$. We generated two types of resource distributions. A "non-surfable" resource (*island resource*), and weakly surfable resource (*drifting resource*).  Both are characterized by a peak in time and space centered at $M_x$ at $m_t$, and $-m_x$ at $\tau - m_t$ (for example, locations 40 and -40 at times 25 and 75, respectively). These pulses have a shared time scale of duration $s_t$ and a spatial scale of extent $s_x$, the standard deviation in the time and space dimension respectively.  The island resource is simply two uncorrelated bivariate normal distributions $$h(x,t) = K\,(\Phi(m_x, m_t, s_x, s_t) + \Phi(-m_x, \tau - m_t, s_x, s_t))$$
where $\Phi$ is the bivariate Gaussian distribution function, and the normalizing constant $K$ is selected such that the average total amount of resource throughout the year is 1, i.e. ${1 \over \tau} \int_T\int_X h(x,t) dx\,dt = 1$.

The drifting resource differs from the island resource in that the total amount of resource at any given time point is 1 ($\int_X h(x,t) dx = 1$).  This property is attained by distributing the resource as a re-scaled beta distribution, where the shape and scale parameters vary sinusoidally in such a way as to make the standard deviations and means match the desired values of $m_x, m_t, s_x, s_t$ (see appendix for details).

Within a given year, the resource is entirely symmetric: $h_y(x,t) = h_y(-x, \tau-t)$.  However, in scenarios exploring climate change we allow the peaks to vary with drift and stochasticity according to:  $m_x(y) \sim {\cal N}(\mu_x + \beta_x\,y, \sigma_x)$ and $m_t(y) \sim {\cal N}(\mu_t + \beta_t\,y, \sigma_t)$, where the $\mu$, $\beta$ and $\sigma$ terms are the mean, slope and variance, respectively, for the location and time duration of the pulse.  Thus, if $\beta=0$ and $\sigma=0$, the conditions are constant across years, if $\beta_x > 0$ there is a drift of the resource towards extremes, if $\beta_t < 0$ there is a shift towards earlier resource pulses.  These trends mirror the pole-ward shift of peak resources and the earlier spring phenology occurring with a warming global climate.  The spatial scales ($s_x$ and $s_t$) remain constant in all of our simulations.

## Metrics

The main metrics we are interested are *migratoriness* and *foraging efficiency*.  The migratoriness is measured as the maximum distance between the median of the population distribution within a given year.  The median is a more reliable metric than the mean, as even as distributions become highly assymetric - pulling the mean in one directon or another - the median can  remain consistently in the middle of the spatial domain indicating a net flux of animals of 0. This is the anticipated (and realized) result of a purely "tactical" model pursuing the periodic resource pulses.

To quantify the foraging efficiency, i.e., the organisms’ ability to track the distribution of the resources over space and time, we use a continuous form of the Bhattacharyya Coefficient [@Bhattacharyya1943] which quantifies the similarity between two distributions.  We compute this coefficient at every time point in a given year, and take the mean across the equilibrium year to determine foraging efficiency (FE). Thus, the foraging efficiency index is: 
$$FE = {1\over \tau} \int_{0}^\tau \int_{0}^{\chi} \sqrt{u(x,t) \, h(x,t)} \,\, dx\,dt$$
where the spatial integral is taken over the domain.  This metric is constrained to be between 0 and 1.

For simulations with a constant resource, we run those until a quasi-equilibrium state is achieved, i.e. where the Bhattacharya index of the population distribution across subsequent years reaches a value of 0.9999.  Once that state is attained, we compute the mirgatoriness and foraging efficiency metrics, and additionally collect the number of years that were required to reach stationarity.

For numerical runs with climate change, we run the simulation for 30 years, since equilibrium is never achieved) and retain the foraging efficiency for the final 10 years, reflecting something like the dynamic long-term average of the system's ability to keep track of climate change.  We also record the migration parameters $\widehat{\theta_y}$ for each year.

## Simulation studies

We explored this model using numerical differencing of a system of ordinary differential equations (ODE's) approximating the PDE in equation (2) with the Runge-Kutte algorithm using the `deSolve` [@Soetaert2010] and `ReacTran` [@Soetaert2012] packages in R.  We additionally used the `nlsLM` function in package `minpack.LM` [@minpack.lm] for robust and fast annual estimation of the migration parameters.  The complete code is available as an R package (`memorymigration`) available on GitHub at https://github.com/EliGurarie/memorymigration and as an interactive Shiny application at https://url.to.be.determined.

We assessed a wide range of parameter values and resource geometries and dynamics with the goal of answering the three main questions:

1. Can this model adapt to a discrete shift in peak resource location and timing?  What is the relative role of memory and sociality?

2. Can this model "acquire" a migratory behavior from a non-migratory initial condition?

3. Can this model adapt when the resource peak shifts in space and time?  What is the role of the reference memory.

# Results

## Adaptation

![](figures/example1_adaptation.png)
**Figure 1. Example of adaptation to a shift in resource peak. The initial (year 0) behavior migrates to locations 50 and -50 at days 15 and 60, whereas the resource peak is at 30 and -30, peaking at times 25 and 75.  (a) shows the first 14 years of the simulation; (b) The centroid of the annual movement of the population is shown in panel b, with dark blue to red colors indicating year 0 to year 40; (c) annual foraging efficiency across years; (d) migration timing parameters for each year, with orange segments indicating arrival and departure from the summering (northern) grounds, and the blue segments indicating timing of arrival and departure at the wintering grounds; (e) migration arrival and departure location across years, with blue and orange indicating winter (southern) and summer (northern) locations.**

We illustrate an example of a population with a particular migration behavior confronted with a major sudden shift in the spatio-temporal distribution of the resource in Figure 1.  In the illustrated scenario, it takes nearly 40 years to attain an equilibrium, and the eventual "optimum" is one where the centroid of the migration lines up exactly with the centroid of the resource, and the arrival timing conincides with the *peak* of resource availability.  Notably, the path to this equilibrium is somewhat indirect, with the summer season, in particular, taking

## Learning to migrate

![](figures/example2_learningtomigrate.png)
**Figure 2. Example of model learning to migrate. The resource is a "weakly drifting" resource and the initial (year 0) condition is non migratory.  The simulation was run for 100 years, and a sampling of those years (labeled) are presented in panel a:  all years from 0 to 10, followerd by 20, 40 and 100.  Otherwise, panels are as in Figure 1.  Parameter values were $\epsilon  = 5$, $\alpha = 500$ and $\beta = 50$.**

Figure 2 illustrates the ability of the model was able to learn to migrate in a weakly drifting resource environment with a narrow pulse of resource peaking at 50 and -50, but a uniform distribution of resource at times 0 and 50.   In order to learn to migrate, the system needed to have a higher exploratory impulse (higher diffusion constant $\epsilon$), a stronger resource advection (higher $\alpha$) and somewhat weaker sociality (lower $\beta$).  The qualitative behavior of this process was to start drifting northward, while slowly developing a weak southward pulse in winter as well.  After first locking in on the summer resource, the southward migration, driven both by high diffusion and high resource following, slowly extended itself until both narrow peaks of resource could be reached consistently.

## Adapting to climate change

# Discussion

We found that the relatively straightforward inclusion of a migration specific memory in a social diffusion-advection framework was sufficient to capture many fundamental phenomena related to migration.  In essense, our model allowed a population of individuals with a hybrid migratory and resource following behavior to adjust the migratory behavior based on recent experiences with the resource location. The model, while simple, was able to emulate the successful navigation of an environment with temporally and spatially isolated seasonal resource patches, the spontaneous emergence of a migratory behavior, and intrinsic robustness to changes in those environmental resources, whether as trends or inter-annual stochasticity.

Our model is somewhat similar to an evolutionary individual-based model [@Anderson2013], which showed that some inherent phenotypic plasticity is required to hedge both against persistent trends and increased stochasticity of resource distributions, but at the risk of sub-optimal access to resources [@Anderson2013].

# References