### 2021-07-09 text/table.Rmd
table

---
title: "Tables"
author: "Elie"
date: "7/9/2021"
output: html_document
---

**Memory migration model**

| Parameter   | | Description  
| ----------- | ----------- | 
| $\alpha$  | Forage following | 
| $\beta$  | Strength of sociality |
| $\lambda$  | Spatial scale of sociality |

**Resource**

**Metrics**