### 2021-07-10 text/table.Rmd
updated shiny + table for MS

---
title: "Tables"
author: "Elie"
date: "7/9/2021"
output: html_document
---

**Memory migration model**

| Parameter   | | Description |
| ----------- | ----------- |
| $\epsilon$  | Diffusion | 
| $\alpha$  | Forage following | 
| $\beta$  | Strength of sociality |
| $\lambda$  | Spatial scale of sociality |
| $\kappa$  | Proportion of reference versus working memory |

**Resource**
| Parameter   | | Description |
| ----------- | ----------- |
| $h$  | Dynamic resource | 
| $\tau$  | Periodicity | 
| $M_x$  | Spatial coordinate of resource peak for the summer season | 
| $-m_x$  | Spatial coordinate of resource peak for the winter season |
| $m_t$  | Timing of resource peak for the summer season | 
| $\tau - m_t$ | Timing of resource peak for the summer season | 
| $s_t$  | Time scale of duration of resource pulse | 
| $s_x$  | Spatial scale of extent of resource pulse |

**Population**
| Parameter   | | Description |
| ----------- | ----------- |
| $t_1$  | Timing of the summer season | 
| $dt_1$  | Duration of the summer season | 
| $t_2$  | Timing of the winter season | 
| $dt_2$  | Duration of the winter season | 
| $x_1$  |  Spatial coordinate of the population centroid for the summer season | 
| $x_2$  |Spatial coordinate of the population centroid for the winter season |

**Metrics**
| Parameter   | | Description |
| ----------- | ----------- |
| $MM_x$  | Spatial migration mismatch |
| $MM_t$  | Temporal migration mismatch |
| $FE$  | Foraging efficiency |
| $SA$  | Spatial adaptation |