### 2021-07-13 text/table.Rmd

---
title: "Tables"
author: "Elie"
date: "7/9/2021"
output: html_document
---

*Memory advection-diffusion PDE*

| Parameter   | Description |
| ----------- | ----------- |
| $\epsilon$  | Diffusion | 
| $\alpha$  | Forage following | 
| $\beta$  | Strength of sociality |
| $\lambda$  | Spatial scale of sociality |
| $\kappa$  | Proportion of reference versus working memory |
| $x_1$, $x_2$  | location of population centroids in summer and winter| 
| $t_1$, $dt_1$  | start and duration of summer season |
| $t_2$, $dt_2$  | start and duration of winter season | 
| $\kappa$ | proportion memory allocated to reference (long-term) memory vs. short-term memory|

*Resource dynamics*

| Parameter   | Description |
| ----------- | ----------- |
| $\tau$  | duration of period (year) | 
| $m_x$, $-m_x$  | spatial coordinate of resource peak for summer and winter | 
| $m_t$, $\tau - m_t$  | timing of resource peak for the summer and winter | 
| $\sigma_x$, $\sigma_t$   | time duration and and spatial scale of resource pulse | 
| $dx$, $dt$  | rate of change of peak locaton and timing of resource | 
| $\psi_x$, $\psi_t$  | standard deviation of peak location and timing |

**Metrics**

| Parameter   | Description |
| ----------- | ----------- |
| $MM_x$  | spatial migration mismatch |
| $MM_t$  | temporal migration mismatch |
| $TE$  | total mismatch |
| $FE$  | foraging efficiency |
| $SA$  | spatial adaptation index |